---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "Viz Wizards - Nick Ferrante, Jeremiah Gaiser, Tanya Evita George, Mrunal Jadhav, Jasdeep Singh Jhajj, Gillian Aine McGinnis, Agastya Deshraju"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

## Description and Motivation:

The data set Meteorite_Landings.csv, which will be used for this project has been sourced from the official website of the United States Government that provides open data. The dimensions of the data set include 10 variables totally which consist of 6 numerical and 4 categorical variables. The reason why this data set was chosen was because of the opportunity to scientifically explore all the factors that go into studying how and why meteorites land on specific parts of the Earth. Using this data set, the project could explain the reasons behind certain historical events and predict the most vulnerable places for meteorite crashes which could help in their scientific study and research in fields such as astrology, data analysis and geology.

Make sure to load the data and use inline code for some of this information.

## Questions

-   What does the distribution of meteor impacts look like across the earth?
    -   Which continents have accumulated the most total mass of meteors?
-   How do historical events relate to the observations and discoveries of meteors?

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
