---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "Viz Wizards - Nick Ferrante, Jeremiah Gaiser, Tanya Evita George, Mrunal Jadhav, Jasdeep Singh Jhajj, Gillian McGinnis, Agastya Deshraju"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
#| echo: false

if (!require("pacman")) 
  install.packages("pacman")
# load packages

pacman::p_load(tidyverse,
               kableExtra) 
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

## Description and Motivation:

The data set Meteorite_Landings.csv, which will be used for this project has been sourced from the official website of the United States Government that provides open data. The dimensions of the data set include 10 variables totally which consist of 6 numerical and 4 categorical variables. The reason why this data set was chosen was because of the opportunity to scientifically explore all the factors that go into studying how and why meteorites land on specific parts of the Earth. Using this data set, the project could explain the reasons behind certain historical events and predict the most vulnerable places for meteorite crashes which could help in their scientific study and research in fields such as astrology, data analysis and geology.

### Metadata

The data includes 10 variables- 6 numerical and 4 categorical. The metadata is shown down below, with a sample of the individual variable as well. The descriptions of what each variable represents is shown as well.

```{r}
#| code-fold: TRUE

meteorite_data <- read.csv("data/Meteorite_Landings.csv")

metadata_meteorite <- tibble(
  Column = names(meteorite_data),
  DataType = sapply(meteorite_data, class),
  SampleData = sapply(meteorite_data, function(column) {
    first_non_na <- column[!is.na(column)][1]
    if (is.numeric(first_non_na)) {
      return(format(first_non_na, nsmall = 2))
    } else {
      return(as.character(first_non_na))
    }
  })
)


metadata_meteorite %>%
  kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F) %>%
  column_spec(1, bold = T)

```

**Description of the metadata**

-   **name**: The name of the meteorite (usually the location of the meteorite landing)
-   **id**: The unique identifier number assigned to the meteorite.
-   **nametype**: Can be one of two different categories:
    -   *valid*: A regular meteorite.
    -   *relict*: A meteorite that has been degraded over the years due to weather.
-   **recclass**: The recommended class of the meteorite; that is classified based on certain characteristics of the meteorite such as, chemical, isotopic, and mineralogical properties.
-   **mass(g)**: The mass of the meteorite, given in grams.
-   **fall**: Can be one of two different categories:
    -   *fell*: Classified as fell when the fall of the meteorite is observed.
    -   *found*: Classified as found when the fall of the meteorite is not observed, but the meteorite was found later.
-   **reclat**: The latitude of the meteorite's landing.
-   **reclong**: The longitude of the meteorite's landing.
-   **GeoLocation**: Combination of the latitude and longitude of the meteorite's landing.

## Questions

-   What does the distribution of meteor impacts look like across the earth?

    -   Which continents have accumulated the most total mass of meteors?

-   How do historical events relate to the observations and discoveries of meteors?

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

## Repository organization

- Main/root: All relevant Quarto markdown code, YAML settings, and folders.
  - `.gitignore`: List of files or file types to not track via version control.
  - `README.md`: Brief project description, viewable for the [repository landing page](https://github.com/INFO-526-S24/project-final-VizWizards).
  - `*.qmd`: Quarto markdown files for the publication of the GitHub website, including the main landing page to contain the final report (`index.qmd`), the project proposal (`proposal.qmd`), a list of project members (`about.qmd`), and the slidedeck for the final presentation (`presentation.qmd`).
- `/data` folder:
  - `Meteorite_Landings.csv`: Raw data file
  - `README.md`: Data dictionary for meteors.csv
- `/images` folder: Any images to be added for the presentation.qmd or GitHub website.
- `/shiny` folder: 
  - `app.R`: All R code for both the Shiny UI and server. To be independent from the main folder to allow for publication on a separate Shiny server.
  - `README.md`: Description of the folder
- Other:
  - `/.github` folder: For automated workflows.
  - `/_extra` folder: For any drafting work or notes not to be included in the final project.
  - `/freeze` folder: Automatically-generated GitHub website publication code.
